Details of specific functions

### 2020-09-07 inst/gitbook/03_specific-tutorials.Rmd
Added between groups tutorial

### `make.metric`

@@

#### Between group metrics

The `group.dist` metric allows to measure the distance between two groups in the multidimensional space.
This function needs to intake several groups and use the option `between.groups = TRUE` in the `dispRity` function.
It calculates the vector normal distance (euclidean) between two groups and returns 0 if that distance is negative.
Note that it is possible to set up which quantiles to consider for calculating the distances between groups.
For example, one might be interested in only considering the 95% CI for each group.
This can be done through the option `probs = c(0.025, 0.975)` that is passed to the `quantile` function.
It is also possible to use this function to measure the distance between the groups centroids by calculating the 50% quantile (`probs = c(0.5)`).

>>>>

### Between groups metrics {#betweengroupmetricsexplain}

One specific category of metrics in the `dispRity` package is the between groups metrics.
As the name suggest, these metrics can be used to calculate the disparity between groups rather than within the groups.
These metrics follow the same classifications as the "normal" (within group) metrics with dimension-level 1, 2 and 3 between groups metrics.
However, at the difference of the "normal" metrics, their input arguments must be `matrix` and `matrix2` (and of course any other additional arguments).
For example, this metric measures the difference in mean between two matrices:

```{r}
## A simple example
mean.difference <- function(matrix, matrix2) {
  mean(matrix) - mean(matrix2)
}
```

You can find the list of implemented between groups metric [here](#betweengroupmetricslist) or design them yourself for your specific needs (potentially using [`make.metric`](#makemetric) for help).

The function works by simply using the two available matrices, with no restriction in terms of dimensions (although you'd probably want both matrices to have the same number of dimensions)

```{r}
## A second matrix
dummy_matrix2 <- matrix(runif(12), 4, 3)

## The difference between groups
mean.difference(dummy_matrix, dummy_matrix2)
```

Beyond this super simple example, it might probably be interesting to use this metric on `dispRity` objects, especially the ones from [`custom.subsets`](#custom-subsets) and [`chrono.subsets`](#chrono-subsets).
In fact, the `dispRity` function allows to apply the between groups metric directly to the `dispRity` objects using the `between.groups = TRUE` option.
For example:

```{r}
## Combining both matrices
big_matrix <- rbind(dummy_matrix, dummy_matrix2)
rownames(big_matrix) <- 1:8

## Making a dispRity object with both groups
grouped_matrix <- custom.subsets(big_matrix,
                                 group = c(list(1:4), list(1:4)))

## Calculating the mean difference between groups
(mean_differences <- dispRity(grouped_matrix,
                              metric = mean.difference,
                              between.groups = TRUE))

## Summarising the object
summary(mean_differences)
## Note how the summary table now indicates
## the number of elements for each group
```

For `dispRity` objects generated by `custom.subsets`, the `dispRity` function will by default apply the metric on the groups in a pairwise fashion.
For example, if the object contains multiple groups, all groups will be compared to each other:

```{r}
## A dispRity object with multiple groups
grouped_matrix <- custom.subsets(big_matrix,
                                 group = c("A" = list(1:4),
                                           "B" = list(1:4),
                                           "C" = list(2:6), 
                                           "D" = list(1:8)))

## Measuring disparity between all groups
summary(dispRity(grouped_matrix, metric = mean.difference,
                 between.groups = TRUE))
```

For `dispRity` objects generated by `chrono.subsets` (not shown here), the `dispRity` function will by default apply the metric on the groups in a serial way (group 1 vs. group 2, group 2 vs. group 3, group 3 vs. group 4, etc...).
However, in both cases (for objects from `custom.subsets` or `chrono.subsets`) it is possible to manually specific the list of pairs of comparisons through their ID numbers:

```{r}
## Measuring disparity between specific groups
summary(dispRity(grouped_matrix, metric = mean.difference,
                 between.groups = list(c(1,3), c(3,1), c(4,1))))

Note that in any case, the order of the comparison can matter.
In our example, it is obvious that `mean(matrix) - mean(matrix2)` is not the same as `mean(matrix2) - mean(matrix)`.

### `make.metric` {#makemetric}

@@

#### Between group metrics {#betweengroupmetricslist}

You can find detailed explanation on how between group metrics work [here](#betweengroupmetricsexplain).
The `group.dist` metric allows to measure the distance between two groups in the multidimensional space.
This function needs to intake several groups and use the option `between.groups = TRUE` in the `dispRity` function.
It calculates the vector normal distance (euclidean) between two groups and returns 0 if that distance is negative.
Note that it is possible to set up which quantiles to consider for calculating the distances between groups.
For example, one might be interested in only considering the 95% CI for each group.
This can be done through the option `probs = c(0.025, 0.975)` that is passed to the `quantile` function.
It is also possible to use this function to measure the distance between the groups centroids by calculating the 50% quantile (`probs = c(0.5)`).
